### 2016-10-07 revision/introduction.tex
minor fixes

The majority of the existing algorithms for OLO are based on two generic
algorithms: \textsc{Follow The Regularizer Leader} (\textsc{FTRL}) and
\textsc{Mirror Descent} (\textsc{MD}). \textsc{FTRL} dates back to the
potential-based forecaster in \cite[Chapter~11]{Cesa-Bianchi-Lugosi-2006} and
its theory was developed in \cite{Shalev-Shwartz-2007}.  The name \textsc{Follow
The Regularized Leader} comes from~\cite{Abernethy-Hazan-Rakhlin-2008}.
Independently, the same algorithm was proposed in~\cite{Nestorov-2009} for
convex optimization under the name \textsc{Dual Averaging} and rediscovered
in~\cite{Xiao-2010} for online convex optimization. Time-varying regularizers
were analyzed in~\cite{Duchi-Hazan-Singer-2011} and generalized in
\cite{Orabona-Crammer-Cesa-Bianchi-2014}. \textsc{MD} was originally proposed in
\cite{Nemirovski-Yudin-1983} and later analyzed in~\cite{Beck-Teboulle-2003} for
convex optimization. In the online learning literature it makes its first
appearance, with a different name, in~\cite{Kivinen-Warmuth-1997}.

Both \textsc{FTRL} and \textsc{MD} are parametrized by a function called
\emph{regularizer}. The function is assumed to be strongly convex with respect
to a given norm. Based on different regularizers different algorithms with
different properties can be instantiated. A summary of algorithms for OLO is
presented in Table~\ref{table:results}.  All of them are instances of
\textsc{FTRL} or \textsc{MD}.

@@

Only one scale-free algorithm based on \textsc{FTRL} was known. It is the
\textsc{AdaHedge}~\cite{de-Rooij-van-Erven-Grunwald-Koolen-2014} algorithm for
learning with expert advice, where the decision set is bounded. An algorithm
based on \textsc{FTRL} that is ``almost'' scale-free is \textsc{AdaGrad
FTRL}~\cite{Duchi-Hazan-Singer-2011}.  This algorithm fail to be scale-free due
to ``off-by-one'' issue; see~\cite{McMahan-2014} and the discussion in
Section~\ref{section:solo-ftrl}. Instead,
\textsc{FTRL-Proximal}~\cite{McMahan-Streeter-2010,McMahan-2014} solves the
off-by-one issue, but it requires proximal regularizers, that in general do not
have a simple form and even in the simple 2-norm case requires bounded domains.

@@

We prove that the regret of \textsc{SOLO FTRL} and \textsc{AdaFTRL} after $T$
rounds is bounded by $O (\sqrt{\sup_{v \in K} f(v) \sum_{t=1}^T\|\ell_t\|_*^2} )$
where $f$ is a non-negative regularizer that is $1$-strongly convex with respect
to a norm $\|\cdot\|$ and $\|\cdot\|_*$ is its dual norm. For \textsc{Scale-Free
MD}, we prove $O (\sqrt{\sup_{u,v \in K} B_f(u,v) \sum_{t=1}^T\|\ell_t\|_*^2} )$
where $B_f$ is the Bregman divergence associated with a $1$-strongly convex
regularizer $f$. In Section~\ref{section:lower-bound}, we show that the
$\sqrt{\sum_{t=1}^T \|\ell_t\|_*^2}$ term in the bounds is necessary by proving
a $\frac{D}{\sqrt{8}} \sqrt{\sum_{t=1}^T\|\ell_t\|_*^2}$ lower bound on the
regret of any algorithm for OLO for any decision set with diameter $D$ with
respect to the primal norm $\|\cdot\|$.

@@

Finally, in Section~\ref{subsection:mirror-descent-lower-bound}, we show
negative results for existing popular variants of \textsc{MD}. We show two
examples of decision sets and sequences of loss vectors of unit norm on which
these variants of \textsc{MD} have $\Omega(T)$ regret.  These results indicate
that \textsc{FTRL} is superior to \textsc{MD}.

>>>>

The majority of the existing algorithms for OLO are based on two generic
algorithms: \textsc{Follow The Regularizer Leader} (\textsc{FTRL}) and
\textsc{Mirror Descent} (\textsc{MD}). \textsc{FTRL} dates back to the
potential-based forecaster in \cite[Chapter~11]{Cesa-Bianchi-Lugosi-2006} and
its theory was developed in \cite{Shalev-Shwartz-2007}.  The name \textsc{Follow
The Regularized Leader} comes from~\cite{Abernethy-Hazan-Rakhlin-2008}.
Independently, the same algorithm was proposed in~\cite{Nestorov-2009} for
convex optimization under the name \textsc{Dual Averaging} and rediscovered
in~\cite{Xiao-2010} for online convex optimization. Time-varying regularizers
were analyzed in~\cite{Duchi-Hazan-Singer-2011} and the analysis tightened in
\cite{Orabona-Crammer-Cesa-Bianchi-2014}. \textsc{MD} was originally proposed in
\cite{Nemirovski-Yudin-1983} and later analyzed in~\cite{Beck-Teboulle-2003} for
convex optimization. In the online learning literature it makes its first
appearance, with a different name, in~\cite{Kivinen-Warmuth-1997}.

Both \textsc{FTRL} and \textsc{MD} are parametrized by a function called
\emph{regularizer}.
%The function is assumed to be strongly convex with respect to a given norm.
Based on different regularizers different algorithms with
different properties can be instantiated. A summary of algorithms for OLO is
presented in Table~\ref{table:results}.  All of them are instances of
\textsc{FTRL} or \textsc{MD}.

@@

Only one scale-free algorithm based on \textsc{FTRL} was known. It is the
\textsc{AdaHedge}~\cite{de-Rooij-van-Erven-Grunwald-Koolen-2014} algorithm for
learning with expert advice, where the decision set is bounded. An algorithm
based on \textsc{FTRL} that is ``almost'' scale-free is \textsc{AdaGrad
FTRL}~\cite{Duchi-Hazan-Singer-2011}.  This algorithm fail to be scale-free due
to ``off-by-one'' issue; see~\cite{McMahan-2014} and the discussion in
Section~\ref{section:solo-ftrl}. Instead,
\textsc{FTRL-Proximal}~\cite{McMahan-Streeter-2010,McMahan-2014} solves the
off-by-one issue, but it requires proximal regularizers. In general, proximal regularizers do not
have a simple form and even the simple 2-norm case requires bounded domains to achieve non-vacuous regret.

@@

We prove that the regret of \textsc{SOLO FTRL} and \textsc{AdaFTRL} on bounded domains after $T$
rounds is bounded by $O (\sqrt{\sup_{v \in K} f(v) \sum_{t=1}^T\|\ell_t\|_*^2} )$
where $f$ is a non-negative regularizer that is $1$-strongly convex with respect
to a norm $\|\cdot\|$ and $\|\cdot\|_*$ is its dual norm. For \textsc{Scale-Free
MD}, we prove $O (\sqrt{\sup_{u,v \in K} B_f(u,v) \sum_{t=1}^T\|\ell_t\|_*^2} )$
where $B_f$ is the Bregman divergence associated with a $1$-strongly convex
regularizer $f$. In Section~\ref{section:lower-bound}, we show that the
$\sqrt{\sum_{t=1}^T \|\ell_t\|_*^2}$ term in the bounds is necessary by proving
a $\frac{D}{\sqrt{8}} \sqrt{\sum_{t=1}^T\|\ell_t\|_*^2}$ lower bound on the
regret of any algorithm for OLO for any decision set with diameter $D$ with
respect to the primal norm $\|\cdot\|$.

@@

Finally, in Section~\ref{subsection:mirror-descent-lower-bound}, we show
negative results for existing popular variants of \textsc{MD}. We show two
examples of decision sets and sequences of loss vectors of unit norm on which
these variants of \textsc{MD} have $\Omega(T)$ regret.  These results indicate
that \textsc{FTRL} is superior to \textsc{MD} in a worst-case sense.

### 2016-10-07 revision/solo-ftrl.tex
minor fixes

In this section, we introduce our first scale-free algorithm; it will be based
on \textsc{FTRL}. The closest algorithm to a scale-free \textsc{FTRL} in the
existing literature is the \textsc{AdaGrad FTRL}
algorithm~\cite{Duchi-Hazan-Singer-2011}. It uses a regularizer on each
coordinate of the form
\begin{equation*}
R_t(w) = R(w) \left(\delta + \sqrt{\sum_{i=1}^{t-1} \|\ell_i\|_*^2} \right).
\end{equation*}
This kind of regularizer would yield a scale-free algorithm \emph{only} for
$\delta=0$.  Unfortunately, the regret bound in~\cite{Duchi-Hazan-Singer-2011}
becomes vacuous for such setting in the unbounded case. In fact, it requires
$\delta$ to be greater than $\|\ell_t\|_*$ for all time steps $t$, requiring
knowledge of the future (see Theorem~5 in~\cite{Duchi-Hazan-Singer-2011}). In
other words, despite of its name, \textsc{AdaGrad FTRL} is not fully adaptive to
the norm of the gradient vectors. Identical considerations hold for the
\textsc{FTRL-Proximal} in \cite{McMahan-Streeter-2010,McMahan-2014}: the
scale-free setting of the learning rate is valid only in the bounded case.

@@

The proof of Theorem~\ref{theorem:regret-solo-ftrl} relies on an inequality
(Lemma~\ref{lemma:useful}).  Related and weaker inequalities
(Lemma~\ref{lemma:sum-of-square-roots-inverses}) were proved
in~\cite{Auer-Cesa-Bianchi-Gentile-2002} and~\cite{Jaksch-Ortner-Auer-2010}. The
main property of this inequality is that on the right-hand side $C$ does
\emph{not} multiply the $\sqrt{\sum_{t=1}^T a_t^2}$ term.

>>>>

In this section, we introduce our first scale-free algorithm; it will be based
on \textsc{FTRL}. The closest algorithm to a scale-free \textsc{FTRL} in the
existing literature is the \textsc{AdaGrad FTRL}
algorithm~\cite{Duchi-Hazan-Singer-2011}. It uses a regularizer on each
coordinate of the form
\begin{equation*}
R_t(w) = R(w) \left(\delta + \sqrt{\sum_{i=1}^{t-1} \|\ell_i\|_*^2} \right).
\end{equation*}
This kind of regularizer would yield a scale-free algorithm \emph{only} for
$\delta=0$.  Unfortunately, the regret bound in~\cite{Duchi-Hazan-Singer-2011}
becomes vacuous for such setting in the unbounded case. In fact, it requires
$\delta$ to be greater than $\|\ell_t\|_*$ for all time steps $t$, requiring
knowledge of the future (see Theorem~5 in~\cite{Duchi-Hazan-Singer-2011}). In
other words, despite of its name, \textsc{AdaGrad FTRL} is not fully adaptive to
the norm of the gradient vectors. Similar considerations hold for the
\textsc{FTRL-Proximal} in \cite{McMahan-Streeter-2010,McMahan-2014}: The
scale-free setting of the learning rate is valid only in the bounded case.

@@

The proof of Theorem~\ref{theorem:regret-solo-ftrl} relies on an inequality
(Lemma~\ref{lemma:useful}).  Related and weaker inequalities, like
Lemma~\ref{lemma:sum-of-square-roots-inverses}, were proved
in~\cite{Auer-Cesa-Bianchi-Gentile-2002} and~\cite{Jaksch-Ortner-Auer-2010}. The
main property of this inequality is that on the right-hand side $C$ does
\emph{not} multiply the $\sqrt{\sum_{t=1}^T a_t^2}$ term.
